---
title: "UK Graduate Outcomes Data Exploration"
output: html_notebook
---

# Introduction

The aim of this document is to use recent graduate employment outcome data to learn more about differences in the types of job and conditions of employment people who graduate with different types of degree enter after graduation

## Data and resource used 

The UK government publishes an official statistics page (13 June 2017) 'Graduate outcomes for all subjects by university'. It's available [here](https://www.gov.uk/government/statistics/graduate-outcomes-for-all-subjects-by-university)

I'm going to download and place three files in the directory `data/`

* [Main Tables: SFR18/2017](https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/619433/SFR18_2017_Institution_subject_table.xlsx)
* [Underlying Data: SFR18/2017](https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/619435/SFR18_2017_Institution_subject_data.csv)
* [Metadata: SFR18/2017](https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/618557/SFR18_2017_metadata.txt)


```{r prep_data}
dir.create("data")

# Main table in excel
download.file(
  "https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/619433/SFR18_2017_Institution_subject_table.xlsx",
  "data/subject_excel.xlsx"
  )

# Underlying data as csv
download.file(
  "https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/619435/SFR18_2017_Institution_subject_data.csv",
  "data/underlying_data.csv"
  )

# metadata as txt
download.file(
  "https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/618557/SFR18_2017_metadata.txt",
  "data/metadata.txt"
  )

```

In this case it was more straightforward to download the files manually, as the versions created using download.file were slightly different and corrupted in the case of the Excel file. First lesson! 

We're going to look mainly at the contents of the file `SFR18_2017_Institution_subject_data.csv`. 

```{r load prereq}
#install.packages("pacman")

pacman::p_load("tidyverse")

```

```{r load_explore_data}

dta <- read_csv("data/SFR18_2017_Institution_subject_data.csv")

```


Now let's think about data tidying

```{r}

dta %>% 
  mutate(
    UKPRN = as.character(UKPRN)
  ) %>% 
  mutate_at(vars(yearsAfterGraduation:prAttIncluded), funs(as.numeric)) %>% 
  filter(!is.na(grads), !is.na(matched)) %>% 
  filter(grads > 0, matched > 0) -> dta2
```

Now let's try to estimate the proportion in employment after 1 to 5 years 

```{r}

dta2 %>% 
  select(subject, yearsAfterGraduation, matched, sustEmpFSorBoth) %>%
  filter(!is.na(sustEmpFSorBoth)) %>% 
  mutate(tmp = matched * sustEmpFSorBoth) %>%

  group_by(yearsAfterGraduation, subject) %>%
  summarise(avg_sustEmpFSorBoth = sum(tmp) / sum(matched)) %>%
  ungroup() %>% 
  mutate(sochealth = case_when(
    subject == "Social studies (excluding economics)" ~ "social",
    TRUE ~ "other"
  ) ) %>% 
  ggplot(aes(x = yearsAfterGraduation, y = avg_sustEmpFSorBoth, group = subject, colour = sochealth))+
  geom_line()

```

Now let's try the same, but with employment only 


```{r}

dta2 %>% 
  select(subject, yearsAfterGraduation, matched, sustEmpOnly) %>%
  filter(!is.na(sustEmpOnly)) %>% 
  mutate(tmp = matched * sustEmpOnly) %>%

  group_by(yearsAfterGraduation, subject) %>%
  summarise(avg_sustEmponly = sum(tmp) / sum(matched)) %>%
  ungroup() %>% 
  mutate(sochealth = case_when(
    subject == "Social studies (excluding economics)" ~ "social",
    TRUE ~ "other"
  ) ) %>% 
  ggplot(aes(x = yearsAfterGraduation, y = avg_sustEmponly, group = subject, colour = sochealth))+
  geom_line()
```

Now median salary

```{r}
dta2 %>% 
  select(subject, yearsAfterGraduation, matched, medianAnnEarn) %>%
  filter(!is.na(medianAnnEarn)) %>% 
  mutate(tmp = matched * medianAnnEarn) %>%

  group_by(yearsAfterGraduation, subject) %>%
  summarise(avg_medianannearn = sum(tmp) / sum(matched)) %>%
  ungroup() %>% 
  mutate(sochealth = case_when(
    subject == "Social studies (excluding economics)" ~ "social",
    TRUE ~ "other"
  ) ) %>% 
  ggplot(aes(x = yearsAfterGraduation, y = avg_medianannearn, group = subject, colour = sochealth))+
  geom_line()
```

Five years after graduation

```{r}

dta2 %>% 
  select(subject, yearsAfterGraduation, matched, medianAnnEarn) %>%
  filter(yearsAfterGraduation == 5) %>% 
  filter(!is.na(medianAnnEarn)) %>% 
  mutate(tmp = matched * medianAnnEarn) %>%
  group_by(yearsAfterGraduation, subject) %>%
  summarise(avg_medianannearn = sum(tmp) / sum(matched)) %>%
  ungroup() %>% 
  ggplot(aes(x = forcats::fct_reorder(subject,avg_medianannearn), y = avg_medianannearn))+
  geom_bar(stat = "identity") + 
  coord_flip()
```


Now to break down by gender

```{r}

dta2 %>% 
  select(subject, sex, yearsAfterGraduation, matched, medianAnnEarn) %>%
  filter(!is.na(medianAnnEarn)) %>% 
  filter(sex != "Female + Male") %>% 
  mutate(tmp = matched * medianAnnEarn) %>%
  group_by(yearsAfterGraduation, subject, sex) %>%
  summarise(avg_medianannearn = sum(tmp) / sum(matched)) %>%
  ungroup() %>% 
  ggplot(aes(x = forcats::fct_reorder(subject,avg_medianannearn), group = sex, fill = sex, y = avg_medianannearn))+
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() +
  facet_wrap(~ yearsAfterGraduation)

```


What is female median salary as proportion of male median salary by institution?

```{r}

dta2 %>% 
  select(subject, sex, yearsAfterGraduation, matched, medianAnnEarn) %>%
  filter(!is.na(medianAnnEarn)) %>% 
  filter(sex != "Female + Male") %>% 
  mutate(tmp = matched * medianAnnEarn) %>%
  group_by(yearsAfterGraduation, subject, sex) %>%
  summarise(avg_medianannearn = sum(tmp) / sum(matched)) %>%
  ungroup() %>% 
  spread(sex, avg_medianannearn) %>% 
  mutate(sex_gap = 100 * (Male - Female) / Male) %>% 
  ggplot(aes(x = forcats::fct_reorder(subject,sex_gap), group = factor(yearsAfterGraduation), fill = factor(yearsAfterGraduation), y = sex_gap))+
  geom_bar(stat = "identity", position = "dodge", colour = "darkgrey") + 
  coord_flip() +
  labs(
    y = "Gender gap in average median graduate earnings (%)",
    x = "University subject",
    fill = "Years after\ngraduation",
    title = "Percentage gap in median pay",
    subtitle = "Gender gap by subject and years after graduation",
    caption = "Data: SFR18/2017 \nhttps://www.gov.uk/government/statistics/graduate-outcomes-for-all-subjects-by-university"
  )  +
  scale_fill_brewer(palette = "Greys", direction = 1) +
  theme_grey() + 
  theme(legend.position = c(0.9, 0.1))

ggsave("figures/gender_gap.png", width = 20, height = 20, units = "cm", dpi = 300)

```


 Let's try to get some summary stats 
 
 
```{r}
dta2 %>% 
  select(subject, sex, yearsAfterGraduation, matched, medianAnnEarn) %>%
  filter(!is.na(medianAnnEarn)) %>% 
  filter(sex != "Female + Male") %>% 
  mutate(tmp = matched * medianAnnEarn) %>%
  group_by(yearsAfterGraduation, sex) %>%
  summarise(avg_medianannearn = sum(tmp) / sum(matched)) %>%
  ungroup() %>% 
  spread(sex, avg_medianannearn) %T>% (
    function(..){
      .. %>% 
        mutate(female_growth = Female / Female[yearsAfterGraduation == 1]) %>% 
        mutate(male_growth = Male / Male[yearsAfterGraduation == 1]) %>% 
        print()
    }
  ) %>% 
  mutate(sex_gap = 100 * (Male - Female) / Male)



```

Want to know the average gender gap in number of graduates by provide and subject 

```{r}

subject_gender_balance <- dta2 %>% 
      group_by(yearsAfterGraduation, providerName, subject) %>% 
      select(yearsAfterGraduation, providerName, subject, sex, grads) %>% 
      filter(sex != "Female + Male") %>% 
      spread(sex, grads) %>% 
      mutate(gender_balance = Male / Female) %>% 
      group_by(yearsAfterGraduation, subject) %>% 
      summarise(median_gender_balance = median(gender_balance)) %>% 
      ungroup() 

subject_gender_balance %>% 
  mutate(graduation_year = case_when(
    yearsAfterGraduation == 1 ~ "12/13",
    yearsAfterGraduation == 3 ~ "10/11",
    yearsAfterGraduation == 5 ~ "08/09"
  )) %>% 
  ggplot(
  aes(x = forcats::fct_reorder(subject,median_gender_balance), y = median_gender_balance,
  fill=factor(graduation_year), group = factor(graduation_year))
  ) +
  geom_bar(stat = "identity", position = "dodge", colour = "darkgrey") + 
  coord_flip() +  
  scale_y_log10(
    breaks = c(0.05, 0.10, 0.25, 0.50, 1, 2, 4, 10, 20),
    labels = c("20F:M", "10F:M", "4F:M", "2F:M", "Equal", "2M:F", "4M:F", "10M:F", "20M:F"),
    limits = c(1/15, 15)
  ) +
  scale_fill_brewer(palette = "Greys", direction = 1) +
  theme_grey() +
    labs(
    y = "Median gender ratio in graduates by subject and year",
    x = "University subject",
    fill = "Graduation Year",
    title = "Gender ratios by subject",
    subtitle = "Gender ratios in study by subject and graduation year",
    caption = "Data: SFR18/2017 \nhttps://www.gov.uk/government/statistics/graduate-outcomes-for-all-subjects-by-university"
  )  +
  theme(legend.position = c(0.85, 0.15))

ggsave("figures/gender_gap_subject.png", width = 20, height = 20, units = "cm", dpi = 300)

        
dta2 %>% 
  select(subject, sex, yearsAfterGraduation, matched, medianAnnEarn) %>%
  filter(!is.na(medianAnnEarn)) %>% 
  filter(sex != "Female + Male") %>% 
  mutate(tmp = matched * medianAnnEarn) %>%
  group_by(yearsAfterGraduation, subject, sex) %>%
  summarise(avg_medianannearn = sum(tmp) / sum(matched)) %>%
  ungroup() %>% 
  spread(sex, avg_medianannearn) %>% 
  mutate(sex_gap = 100 * (Male - Female) / Male) %>% 
  inner_join(
    subject_gender_balance
  ) %>% 
  select(yearsAfterGraduation, subject, median_gender_balance, sex_gap) %>% 
  ggplot(aes(x = median_gender_balance, y = sex_gap)) + 
  scale_x_log10(
    breaks = c(0.05, 0.10, 0.25, 0.50, 1, 2, 4, 10, 20),
    labels = c("20F:M", "10F:M", "4F:M", "2F:M", "Equal", "2M:F", "4M:F", "10M:F", "20M:F")
                ) + 
  geom_point() +
  facet_wrap(~yearsAfterGraduation) +
  geom_smooth( col = "black")
  

```


Let's now try to get a sense of how student ratios have changed over time overall


```{r}
subject_gender_balance <- dta2 %>% 
      group_by(yearsAfterGraduation, providerName, subject) %>% 
      select(yearsAfterGraduation, providerName, subject, sex, grads) %>% 
      filter(sex != "Female + Male") %>% 
      spread(sex, grads) %>% 
      mutate(gender_balance = Male / Female) %>% 
      group_by(yearsAfterGraduation, subject) %>% 
      summarise(median_gender_balance = median(gender_balance)) %>% 
      ungroup() 

subject_gender_balance %>% 
  mutate(females_per_males = 1 / median_gender_balance) %>% 
  mutate(graduation_year = case_when(
    yearsAfterGraduation == 1 ~ "12/13",
    yearsAfterGraduation == 3 ~ "10/11",
    yearsAfterGraduation == 5 ~ "08/09"
  )) %>% View()


```