---
title: "RCUK Funding"
output: html_notebook
---

# Introduction 

This document will explore the distribution of research done by council in the UK, and how it has changed over time.

 Links 
http://www.rcuk.ac.uk/research/efficiency/successrates/


Subjects by research council are described here: 

Arts and Humanities Research Council (AHRC)
The AHRC supports research across a wide subject domains including:

history (ancient, medieval and modern);
classics;
archaeology;
modern languages and linguistics;
English language and literature;
the visual arts and media;
librarianship, information and museum studies;
philosophy, law, religious studies;
music and creative and performing arts.
Biotechnology and Biological Sciences Research Council (BBSRC)
BBSRC funds research in the biosciences, in areas including:

Genomics, stem cell biology, and bio-nanotechnology, that provide a basis for new technologies in healthcare, food safety, plant and livestock breeding, and bio-processing;
Whole organism biology relevant to the understanding of diet and health, ageing, animal health and welfare, infectious diseases and immunity, and crop productivity;
Biological populations and systems that underpin agricultural sustainability, biodiversity and novel bio-based and renewable processes for energy and manufacturing.

Engineering and Physical Sciences Research Council (EPSRC)
EPSRC supports fundamental research in the physical sciences through to more applied topics in engineering and technology. This includes research in the following areas:
mathematics;
chemistry;
physics;
materials science;
engineering;
computer science, including high performance computing;
energy research;
research into the built environment;
information and communications technology;
research into innovative manufacturing.
Economic and Social Research Council (ESRC)
ESRC funds social science and economic research across a wide of disciplines encompassing:

Demography and human geography
Development studies
Economics
History
Education
Environmental planning
Linguistics
Management and business studies
Political science and international studies
Psychology
Social anthropology
Social policy
Social work
Sociology
Law and legal studies
Tools, technologies and methods


Medical Research Council (MRC)
The MRC’s remit covers the full range of medical research from studies of molecules to the implementation of research findings into clinical practice. This includes such topics as:
the use of animals in research;
antibiotic resistance;
brain sciences;
genomics and proteomics;
health of the public;
intensive care;
patient safety research;
stem cell research.


Natural Environment Research Council (NERC)
NERC funds environmental research, survey and observation work across a wide spectrum of disciplines, including:
the geo- and earth sciences, hydrology, soil science, atmospheric research and oceanography;
biological and microbiological research on animal and plant biodiversity, population dynamics and ecology;
climate change research;
environmental chemistry and physics;
satellite based Earth observation;
polar research;
management of land and natural resources.

Science and Technology Facilities Council (STFC)
Amongst others, STFC's areas of research include:
astronomy;
computational science;
energy;
nuclear physics;
particle physics;
space science.

Source: http://www.rcuk.ac.uk/research/Areas/


# Regional funding data 

http://www.rcuk.ac.uk/about/aboutrcs/research-funding-across-the-uk/



# Page 
 - success rates for grant funding 
 http://www.rcuk.ac.uk/research/efficiency/successrates/
 
 Provides separate links to 
 
http://www.ahrc.ac.uk/about/statistics/competitionstatistics/
http://www.bbsrc.ac.uk/funding/apply/post_award_overview.html
http://www.epsrc.ac.uk/funding/successrates/Pages/successrates.aspx
http://www.esrc.ac.uk/about-us/strategy-and-priorities/demand-management/funding-decisions
http://www.mrc.ac.uk/documents/xls-csv/success-rates/
http://www.nerc.ac.uk/funding/application/howtoapply/success/
http://www.stfc.ac.uk/funding/research-grants/

Not at all standardized

Closest to usable infromation from THES article, 2016
https://www.timeshighereducation.com/news/uk-grant-success-rates-prompt-worldwide-comparisons

Table : overall UK research council success rates 

```{r}
require(tidyverse)

overall_by_council <- tribble(
  ~council_name, ~n_applications, ~n_grants, ~success_pct, ~amount, 
  "Arts & Humanities Research Council", 246, 68, 28, 29851958,
  "Biotechnology & Biological Sciences Research Council", 1621, 392, 24, 178719000,
  "Engineering & Physical Sciences Research Council", 2419, 786, 33, 509476006,
  "Economic & Social Research Council", 429, 52, 12, 24107303,
  "Medical Research Council", 1993, 402, 20, 283520000,
  "Natural Environment Research Council", 1036, 311, 30, 86871071
)
# For year 2016 
```
Now to plot this
```{r}
overall_by_council %>% 
  ggplot(aes(x = forcats::fct_reorder(council_name, amount), y = amount)) + 
  geom_bar(stat = "identity") +
  coord_flip()+ 
  labs(x = )
```

Let's try a treemap

```{r}
require(treemap)
treemap(
  overall_by_council,
  index = "council_name",
  vSize = "amount",
  title ="Research funding by council in 2016"
  
)

```
 Now research grant success rates 
 
```{r}
overall_by_council %>% 
  mutate(success_rate = n_grants / n_applications) %>%  
  treemap(
    .,
    index = "council_name",
    vSize = "success_rate",
    title ="Grant success rates by council in 2016"
  
)
 
```

Now let's look at the number of students who enter each subject 

We have this data from the following location:

https://www.hesa.ac.uk/news/11-01-2018/sfr247-higher-education-student-statistics/subjects

```{r}
read_csv(
  "data/sfr247-figure-14 (1).csv",
  skip = 13, col_names = F 
  ) -> tmp

tmp[1,1] <- "year"
tmp[1,2:5] <- 2012
tmp[1,6:9] <- 2013
tmp[1, 10:13] <- 2014
tmp[1, 14:17] <- 2015
tmp[1, 18:21] <- 2016
tmp
tmp2 <- tmp[-1,]
names(tmp2) <- tmp[1,]
tmp2

dta_tidy <- bind_rows(
  tmp %>% (function(..){
    .. %>% 
    select(X1:X5) %>% mutate(year = 2012) %>% slice(-c(1,2)) -> ..
    names(..) <- c("subject", "female", "male", "other", "total", "year")
    ..
  }),
  tmp %>% (function(..){
    .. %>% 
      select(X1, X6:X9) %>% mutate(year = 2013) %>% slice(-c(1,2)) -> ..
    names(..) <- c("subject", "female", "male", "other", "total", "year")
    ..
  }),
  tmp %>% (function(..){
    .. %>% 
      select(X1, X10:X13) %>% mutate(year = 2014) %>% slice(-c(1,2)) -> ..
    names(..) <- c("subject", "female", "male", "other", "total", "year")
    ..
  }),
  tmp %>% (function(..){
    .. %>% 
      select(X1, X10:X13) %>% mutate(year = 2014) %>% slice(-c(1,2)) -> ..
    names(..) <- c("subject", "female", "male", "other", "total", "year")
    ..
  }),
  tmp %>% (function(..){
    .. %>% 
      select(X1, X14:X17) %>% mutate(year = 2015) %>% slice(-c(1,2)) -> ..
    names(..) <- c("subject", "female", "male", "other", "total", "year")
    ..
  }),
    tmp %>% (function(..){
    .. %>% 
      select(X1, X18:X21) %>% mutate(year = 2016) %>% slice(-c(1,2)) -> ..
    names(..) <- c("subject", "female", "male", "other", "total", "year")
    ..
  })
) %>% 
  gather(female:total, key = gender, value = "number") %>% 
  mutate(number = number %>% 
           stringr::str_replace(",", "") %>%
           as.numeric
  )


```

Now to get a lookup for students to research council


```{r}

subject_council_lookup <- tribble(
  ~subject_name, ~council,
"(1) Medicine & dentistry", "mrc",
"(2) Subjects allied to medicine", "mrc",
"(3) Biological sciences", "bbsrc",
"(4) Veterinary science", "nerc",
"(5) Agriculture & related subjects", "nerc",
"(6) Physical sciences", "epsrc",
"(7) Mathematical sciences", "epsrc",
"(8) Computer science", "epsrc",
"(9) Engineering & technology", "epsrc",
"(A) Architecture, building & planning", "esrc",
"(B) Social studies", "esrc",
"(C) Law", "esrc",
"(D) Business & administrative studies", "esrc",
"(E) Mass communications & documentation", "esrc",
"(F) Languages", "ahrc",
"(G) Historical & philosophical studies", "arhc",
"(H) Creative arts & design", "ahrc",
"(I) Education", "esrc",
"(J) Combined", NA
)

council_abbreviation_lookup <- tribble(
  ~abbreviation, ~full_name,
  "mrc","Medical Research Council",
  "esrc", "Economic & Social Research Council",
  "epsrc","Engineering & Physical Sciences Research Council",
  "nerc","Natural Environment Research Council",
  "ahrc", "Arts & Humanities Research Council",
  "bbsrc","Biotechnology & Biological Sciences Research Council"
)


```
Now let's try to join the data 
```{r}
subject_council_lookup %>% 
  inner_join(
    dta_tidy %>% 
      filter(year == 2016) %>% 
      filter(gender == "total"),
    by = c("subject_name"="subject")
    ) %>% 
  group_by(council) %>% 
  summarise(number = sum(number, na.rm = T)) %>% 
  filter(!is.na(council)) %>% 
  inner_join(
    council_abbreviation_lookup,
    by = c("council"="abbreviation")
  ) %>% 
  treemap(
    .,
    index = "full_name",
    vSize = "number",
    title ="Approximate number of students by research council area"
    
  )
    

```

Now let's look at association

```{r}

subject_council_lookup %>% 
  inner_join(
    dta_tidy %>% 
      filter(year == 2016) %>% 
      filter(gender == "total"),
    by = c("subject_name"="subject")
    ) %>% 
  group_by(council) %>% 
  summarise(number = sum(number, na.rm = T)) %>% 
  filter(!is.na(council)) %>% 
  inner_join(
    council_abbreviation_lookup,
    by = c("council"="abbreviation")
  ) %>% 
  left_join(
    overall_by_council %>% select(council_name, research_funding = amount), 
    by = c("full_name"="council_name")
  ) %>% 
  select(
    council_name = full_name,
    number_of_students = number, 
    research_funding
  ) %>% 
  left_join(
    council_abbreviation_lookup,
    by = c("council_name"="full_name")
  ) %>% 
  mutate(research_funding = research_funding / 1000000) %>% 
  mutate(number_of_students = number_of_students / 1000) %>% 
  ggplot(aes(x = number_of_students, y = research_funding, label = abbreviation)) + 
  geom_label() +
  labs(y = "Research council funding  allocated in 2016 (£Millions)",
         x = "Number of students enrolling in 2016 (thousands)",
       title = "Differences between research fields: students vs research funds"
  )

```

Subjects by research council are described here: 

Arts and Humanities Research Council (AHRC)
The AHRC supports research across a wide subject domains including:

history (ancient, medieval and modern);
classics;
archaeology;
modern languages and linguistics;
English language and literature;
the visual arts and media;
librarianship, information and museum studies;
philosophy, law, religious studies;
music and creative and performing arts.

Biotechnology and Biological Sciences Research Council (BBSRC)
BBSRC funds research in the biosciences, in areas including:
Genomics, stem cell biology, and bio-nanotechnology, that provide a basis for new technologies in healthcare, food safety, plant and livestock breeding, and bio-processing;
Whole organism biology relevant to the understanding of diet and health, ageing, animal health and welfare, infectious diseases and immunity, and crop productivity;
Biological populations and systems that underpin agricultural sustainability, biodiversity and novel bio-based and renewable processes for energy and manufacturing.

Engineering and Physical Sciences Research Council (EPSRC)
EPSRC supports fundamental research in the physical sciences through to more applied topics in engineering and technology. This includes research in the following areas:
mathematics;
chemistry;
physics;
materials science;
engineering;
computer science, including high performance computing;
energy research;
research into the built environment;
information and communications technology;
research into innovative manufacturing.

Economic and Social Research Council (ESRC)
ESRC funds social science and economic research across a wide of disciplines encompassing:
Demography and human geography
Development studies
Economics
History
Education
Environmental planning
Linguistics
Management and business studies
Political science and international studies
Psychology
Social anthropology
Social policy
Social work
Sociology
Law and legal studies
Tools, technologies and methods


Medical Research Council (MRC)
The MRC’s remit covers the full range of medical research from studies of molecules to the implementation of research findings into clinical practice. This includes such topics as:
the use of animals in research;
antibiotic resistance;
brain sciences;
genomics and proteomics;
health of the public;
intensive care;
patient safety research;
stem cell research.


Natural Environment Research Council (NERC)
NERC funds environmental research, survey and observation work across a wide spectrum of disciplines, including:
the geo- and earth sciences, hydrology, soil science, atmospheric research and oceanography;
biological and microbiological research on animal and plant biodiversity, population dynamics and ecology;
climate change research;
environmental chemistry and physics;
satellite based Earth observation;
polar research;
management of land and natural resources.

Science and Technology Facilities Council (STFC)
Amongst others, STFC's areas of research include:
astronomy;
computational science;
energy;
nuclear physics;
particle physics;
space science.

Source: http://www.rcuk.ac.uk/research/Areas/